Random Forests, LAs

### 2023-11-21 src/la_RF.Rmd
CAST

## RF spatial with CAST

@@

```{r}
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

@@

@@

@@

@@

@@

@@

@@

rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(train.region, test.region, Rsquared) %>% 
  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  theme(legend.position = "none")
  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

>>>>

## RF with CAST

@@

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 11) # 9 after adding South

@@

print(model.all)
# Random Forest 
# 
# 5984 samples
#   10 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 4718, 4690, 5040, 4718, 5055, 4704, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE        Rsquared   MAE       
#    2    variance    0.02194750  0.9015001  0.01162254
#    2    extratrees  0.02245492  0.8984940  0.01229071
#    6    variance    0.02208770  0.9025087  0.01175267
#    6    extratrees  0.02165214  0.9081808  0.01161281
#   10    variance    0.02447416  0.8859265  0.01326348
#   10    extratrees  0.02204464  0.9064730  0.01184945
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 6, splitrule = extratrees and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.slag", "spatial lag of web density",
                                               ifelse(rowname == "n.lag", "temporal lag of web density",
                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag  of web density",
                                                             ifelse(rowname == "n.nearest.city", "web density in nearest city", rowname)))))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/varimp_LA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)

@@

### train the model in a loop per region

@@

stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken
```

@@

### Test on all other regions

@@

```{r}
# train on one region, test on all the rest

@@

path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(train.region, test.region, Rsquared) %>% 
  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  labs(y = "", x = "") +
  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
  theme(#legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
ggsave(path.out)

  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
